---
title: "Lab2_DivaMC"
output: html_document
date: "2023-02-25"
---


### **Part 1: Introduction to modeling using basic R syntax**

_Create a variable puppies equal to the number of puppies you’d like to have._
```{r echo=TRUE}
 puppies<-22
 puppies
```

_Create a variable puppy_price, which is how much you think a puppy costs._
```{r echo=TRUE}
 puppy_price<-100
 puppy_price

```

_Create a variable total_cost that has the total cost of all of your puppies._
```{r echo=TRUE}
 total_cost <- puppies * puppy_price
 total_cost
```

_Create a boolean variable too_expensive, set to TRUE if the cost is greater than $1,000._
```{r echo=TRUE}
 too_expensive <- total_cost >= 1000
 too_expensive
```

_Create a variable max_puppies, which is the number of puppies you can afford_ for $1,000._
```{r echo=TRUE}
 max_puppies<-10
 max_puppies
```


### **Part 2: Manipulating variables and learning how to use new functions**

_Assign your name to the variable my_name in four different ways._
```{r echo=TRUE}
 my_name <- "Diva M Camp"
 my_name = "Diva M Camp"
 my_name -> "Diva M Camp"
 assign(my_name, "Diva M Camp")
 my_name

 "Diva M Camp" -> my_name
 my_name

 rm("Diva M Camp")
```

_Assign your height (in inches) to a variable my_height._
```{r echo=TRUE}
 my_height <- 64.5
 my_height
```

_Assign your favorite day to a varialbe favorite_day as a date object._
```{r echo=TRUE}
 favorite_day <- "Halloween"
 favorite_day
```

_Assign your favorite quote to a variable favorite_quote. How many characters does your favorite quote have?_
```{r echo=TRUE}
 favorite_quote <- "Are you a Mexi-can or a Mexi-can't?"
 favorite_quote

 nchar(favorite_quote)

```

_Show what type of objects my_name, my_height, favorite_day, and favorite_quote are._
```{r echo=TRUE}
 class(my_name)
 class(my_height)
 class(favorite_day)
 class(favorite_quote)
```

_Coerce these variables to numerics and describe what happens._
```{r echo=TRUE}
 as.numeric(my_name)
 as.numeric(my_height)
 as.numeric(favorite_day)
 as.numeric(favorite_quote)
```

_Create a vector variable named id that contains my_name, my_height, favorite_day, and favorite_quote._
```{r echo=TRUE}
 id <- c(my_name, my_height, favorite_day, favorite_quote)
 id
```

_What class is id? Did the classes for my_name, my_height, favorite_day, and favorite_quote change when they were stored in id? Did the classes change for the variables themselves?_
```{r echo=TRUE}
 class(id)
```

The numeric variables became character variables, because all variables in the vector need to be the same class.

_Try using cat and paste with id as a function argument. How do the results differ? What happens when we use cat and paste at the same time (i.e. f(g(x)))? What happens if we change the order we use them (i.e. g(f(x)))?_
```{r echo=TRUE}
 cat(id)
 paste(id)
 cat(paste(id))
 paste(cat(id))
 paste(my_name, my_height, favorite_day, favorite_quote)
```

Cat concatenates each variable in the vector into a single string. Paste just seems to print the vector as stored. cat(paste(id)) seems to concatenate the variables onto a string. Whereas paste(cat(id)) concatenates the string variables, then prints character(0). 

_How would you determine the difference between cat and paste using R documentation (from within RStudio)? What is a great internet resource to use as discussed in the book?_

Click on the help tab on the lower right quadrant, and type the function we're looking for. Or type onto the console. 

```{r echo=TRUE}
 ?cat
 ?paste
```

_What do sep and collapse arguments for paste do? If we wanted to append each character variable in our vector id with a new line (i.e. "\n") would we use sep or collapse?_

Sep separates the string variables by the character(s) determined between "". Collapse joins with a determined character string. 

_Display the contents of id using a combination of cat and paste with the appropriate arguments for paste._
```{r echo=TRUE}
 cat(paste(id, collapse=" / "))
```


### **Part 3: Accessing data in GitHub and mastering order of operations**

_Please calculate a (using min), b (using max), and c (using mean) from data. Pay careful attention to follow order of operations (PEMDAS)_
```{r echo=TRUE}
load("~/student2/data.RData")

a <- min(data)
a
b <- max(data)
b
c <- mean(data)
c

 x_positive <- ((-b)+sqrt((b*b)-(4*a*c)))/(2*a)
 x_positive
 x_negative <- ((-b)-sqrt((b*b)-(4*a*c)))/(2*a)
 x_negative
```

_Use a, b, and c to solve for x using the quadratic equation. You will need to do this for + and - in the quadratic equation (i.e. you should have TWO answers). Please display the answer as x = {x1, x2} using cat and paste._

```{r echo=TRUE}
x <- c(x_positive, x_negative)
x

cat(paste(x))
```
